Basic Model formulation

### 2015-02-04 documentation/model_formulation.Rmd
update model formulation, produce pdf

---
title: "Basic Model formulation"
author: "Florian D. Schneider"
date: "Tuesday, November 18, 2014"
output:
  pdf_document:
    fig_height: 4
    fig_width: 6
  html_document:
    keep_md: yes
fontsize: 12pt
geometry: margin=1.4in
csl: ../manuscript/amnat.csl
bibliography: ../manuscript/cas02.bib
---

@@

@@

for the set of parameters given above, we get the following dynamics

@@

@@

@@

@@

>>>>

---
title: "Basic Model formulation"
author: "Florian D. Schneider"
date: "Tuesday, February 4th, 2014"
output:
  pdf_document:
    fig_height: 4
    fig_width: 6
  html_document:
    keep_md: yes
fontsize: 12pt
geometry: margin=1.4in
csl: ../manuscript/amnat.csl
bibliography: ../manuscript/cas02.bib
---

@@

```{r, eval = FALSE, echo = FALSE, warning = FALSE}
animateCA(spexmodel, "ca_spexmodel.gif")

@@

for the set of parameters given above, we can project the following dynamics from solving the ODE (CA model simulation as grey line).

@@

The following shows bifurcation diagrams over an environmental gradient.

@@

The following shows bifurcation diagrams over a livestock density gradient.

@@

A dynamic Shiny implementation of that can be found under 'documentation/interactive_spexmodel_bifurcations.Rmd' ( run in RStudio, specify parallel backend.).

The simulation backend is now all set up. As a final step, we want to check out if the model can be solved analytically.

@@

I tried to solve the equilibria of the equation system by assuming

$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 0 $$

$$  0 = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$

and then solving towards $\rho_{11}$. The simplification process can be found in the file 'analytical.Rmd'. It  leads to 4th order polynomials, which makes it impossible to solve. 
However, solutions might be possible for certain cases, e.g. with only facilitation, or only associational protection, and all other effects set to zero.

For instance, assuming competition $c = 0$ and attraction $v = 0$ yields a much simpler equation of second order

$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK) \rho_1^{3}- (-2br-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+2K-amhK-aLK- brK) \rho_1^{2}   + (-mK+aLK-br+3amhK+2farhK+2brahK-2fr+brah-2frK+2aLK) \rho_1^{3}-  (2farh-amhK-aLK-2farhK-brahK+2fr+br)\rho_1^{4} + (2farh + brah)\rho_1^{5}  )     \rho_{11}  + (aLpK\rho_1^{-1}+ 2aLpK  + aLpK + (-frK-brfK-3aLpK+2vLK) \rho_1  + (frbrf+frK+brfK-rahfK-bfrahK+aLpK) \rho_1^{2} + (rahf +bfrah+rahfK +bfrahK-fr -brf)\rho_1^{3} + (-rahf-bfrah) \rho_1^{4}    )   \rho_{11}^2   $$
